Translation Elicitation: laboratory notes

### 2022-11-21 index.qmd
Add questionnaire analyses and restructure the targets script.

```{r}
#| label: tbl-stimuli
#| tbl-cap: "Lexical frequency, phonological neigbourhood size, and levenshtein distance of the stimuli words."
#| tbl-subcap: "Median and standard deviation (*SD*) of lexical frequencies of target words (expressed in counts per million and as Zipf scores), density of higher-frequency phonological neighbourhoods of target words, and normalised Levenshtein phonological similarity between the presented and the target words."
stimuli %>% 
    select(group, word_1, freq_2, freq_zipf_2, nd, lv) %>% 
    group_by(group) %>% 
    summarise(
        across(
            freq_2:lv, 
            list(
                median = ~median(., na.rm = TRUE),
                sd = ~sd(., na.rm = TRUE)
            )
        )
    ) %>%
    gt(rowname_col = "group") %>% 
    fmt_number(freq_2_median:nd_sd) %>%
    fmt_percent(matches("lv|global")) %>% 
    tab_spanner(
        md("**Freq. (million)**"),
        c("freq_2_median", "freq_2_sd", "freq_zipf_2_median", "freq_zipf_2_sd")
    ) %>% 
    tab_spanner(md("**Freq. (Zipf)**"), matches("freq_zipf_2")) %>% 
    tab_spanner(md("**Neighbours**"), matches("nd")) %>% 
    tab_spanner(md("**Levenshtein**"), matches("lv")) %>% 
    cols_label(
        freq_2_median = md("Median"),
        freq_2_sd = md("*SD*"),
        freq_zipf_2_median = md("Median"),
        freq_zipf_2_sd = md("*SD*"),
        nd_median = md("Median"),
        nd_sd = md("*SD*"),
        lv_median = md("Median"),
        lv_sd = md("*SD*")
    ) %>% 
    summary_rows(
        columns = c(matches("median"), -matches("lv")),
        fns = list(Median = "median", SD = "sd")
    ) %>% 
    summary_rows(
        columns = matches("lv"),
        fns = list(Median = "median", SD = "sd"), 
        formatter = fmt_percent
    ) %>% 
    tab_style(
        style = cell_text(align = "left", weight = "bold", size = "medium"),
        locations = cells_title(groups = "title")
    ) %>% 
    tab_style(
        style = cell_text(align = "left", size = "medium", style = "italic"),
        locations = cells_title(groups = "subtitle")
    ) %>% 
    tab_style(
        style = cell_text(align = "left", size = "medium"),
        locations = cells_source_notes()
    ) %>% 
    tab_style(
        style = cell_text(align = "left", size = "medium"),
        locations = cells_stub_grand_summary()
    ) %>% 
    tab_style(
        style = cell_borders(sides = c("left", "right"), color = "white"),
        locations = list(
            cells_body(),
            cells_row_groups(),
            cells_column_spanners(), 
            cells_stub(),
            cells_stub_grand_summary(), 
            cells_source_notes(),
            cells_column_labels(),
            cells_stubhead()
        )
    ) %>% 
    tab_style(
        style = list(
            cell_text(align = "left", size = "medium"),
            cell_borders(sides = "all", color = "white")
        ),
        locations = cells_stub()
    ) %>% 
    tab_style(
        style = cell_text(
            align = "center", 
            size = "medium",
            weight = "normal",
            style = "normal"
        ),
        locations = list(
            cells_column_spanners(),
            cells_column_labels()
        )
    ) %>% 
    tab_style(
        style = cell_borders(sides = "all", color = "white"),
        locations = cells_body(columns = 1:9)
    ) %>% 
    tab_style(
        style = cell_borders(sides = "top", color = "white"),
        locations = cells_body(columns = 1:9)
    )
```

@@

```{r}
#| label: fig-stimuli-freq
stimuli %>% 
    distinct(group, word_1, .keep_all = TRUE) %>% 
    left_join(count(., group, name = "n_total")) %>% 
    mutate(test_language = paste0(group,  " (N = ", n_total, ")")) %>% 
    ggplot(aes(freq_zipf_2, fill = group)) +
    facet_wrap(~group, ncol = 1) +
    geom_histogram(bins = 25, colour = "white") +
    labs(
        x = "Lexical frequency (Zipf score)\nExtracted from CLEARPOND database", 
        y = "# trials", 
        fill = "Group"
    )  +
    scale_y_continuous(breaks = seq(0, 12, 2)) +
    theme(
        legend.position = "none",
        panel.grid.major.y = element_line(colour = "grey", linetype = "dotted")
    )

@@

```{r}
#| label: model-rocs
model_rocs %>% 
    bind_rows(.id = "model") %>% 
    ggplot(aes(1-specificity, sensitivity, colour = model)) +
    facet_wrap(~model) +
    geom_line(aes(group = .draw), alpha = 0.1) +
    labs(x = "1 - Specificity", y = "Sensitivity", colour = "Model") + 
    scale_y_continuous(labels = percent) +
    scale_x_continuous(labels = percent) +
    theme_custom()

@@

```{r}
#| label: tbl-model-coefs
coefs <- fixef(fit_4) %>% 
    as.data.frame() %>% 
    rownames_to_column("variable") %>% 
    clean_names() %>% 
    mutate(across(estimate:q97_5, ~ifelse(variable=="Intercept", inv_logit_scaled(.), ./4)))

str_repl <- c(
    "Intercept" = "Intercept",
    "freq_zipf_2_std" = "Frequency [+1 SD]",
    "nd_std" = "Neighbours [+1 SD]",
    "lv_std" = "Levenshtein [+1 SD]",
    "nd_std:lv_std" = "Neigbours \u00d7 Cognateness"
)

@@

coefs %>% 
    mutate(
        variable = variable %>% 
            str_remove("b_") %>% 
            factor(levels = names(str_repl), labels = str_repl)
    ) %>% 
    select(-est_error) %>% 
    gt() %>% 
    fmt_percent(2:4) %>% 
    cols_merge(c("q2_5", "q97_5"), pattern = "[{1}, {2}]") %>% 
    cols_label(
        variable = md("**Predictor**"),
        estimate = md("**Mean**"),
        q2_5 = md("**95\\% CrI**"),
    ) %>% 
    tab_source_note(tab_caption) %>% 
    cols_align(align = "left")
```

@@

@@

>>>>

```{r}
#| label: tbl-stimuli
#| tbl-cap: "Lexical frequency, phonological neigbourhood size, and levenshtein distance of the stimuli words."
#| tbl-subcap: "Median and standard deviation (*SD*) of lexical frequencies of target words (expressed in counts per million and as Zipf scores), density of higher-frequency phonological neighbourhoods of target words, and normalised Levenshtein phonological similarity between the presented and the target words."
stimuli %>% 
    select(group, word_1, freq_2, freq_zipf_2, nd, lv) %>% 
    group_by(group) %>% 
    summarise(
        across(
            freq_2:lv, 
            list(
                median = ~median(., na.rm = TRUE),
                sd = ~sd(., na.rm = TRUE)
            )
        )
    ) %>%
    gt(rowname_col = "group") %>% 
    fmt_integer(n) %>% 
    fmt_number(freq_2_median:nd_sd) %>%
    fmt_percent(matches("lv|global")) %>% 
    tab_spanner(
        md("**Freq. (million)**"),
        c("freq_2_median", "freq_2_sd", "freq_zipf_2_median", "freq_zipf_2_sd")
    ) %>% 
    tab_spanner(md("**Freq. (Zipf)**"), matches("freq_zipf_2")) %>% 
    tab_spanner(md("**Neighbours**"), matches("nd")) %>% 
    tab_spanner(md("**Levenshtein**"), matches("lv")) %>% 
    cols_label(
        freq_2_median = md("Median"),
        freq_2_sd = md("*SD*"),
        freq_zipf_2_median = md("Median"),
        freq_zipf_2_sd = md("*SD*"),
        nd_median = md("Median"),
        nd_sd = md("*SD*"),
        lv_median = md("Median"),
        lv_sd = md("*SD*")
    ) %>% 
    summary_rows(
        columns = c(matches("median"), -matches("lv")),
        fns = list(Median = "median", SD = "sd")
    ) %>% 
    summary_rows(
        columns = matches("lv"),
        fns = list(Median = "median", SD = "sd"), 
        formatter = fmt_percent
    ) %>% 
    tab_style(
        style = cell_text(align = "left", weight = "bold", size = "medium"),
        locations = cells_title(groups = "title")
    ) %>% 
    tab_style(
        style = cell_text(align = "left", size = "medium", style = "italic"),
        locations = cells_title(groups = "subtitle")
    ) %>% 
    tab_style(
        style = cell_text(align = "left", size = "medium"),
        locations = cells_source_notes()
    ) %>% 
    tab_style(
        style = cell_text(align = "left", size = "medium"),
        locations = cells_stub_grand_summary()
    ) %>% 
    tab_style(
        style = cell_borders(sides = c("left", "right"), color = "white"),
        locations = list(
            cells_body(),
            cells_row_groups(),
            cells_column_spanners(), 
            cells_stub(),
            cells_stub_grand_summary(), 
            cells_source_notes(),
            cells_column_labels(),
            cells_stubhead()
        )
    ) %>% 
    tab_style(
        style = list(
            cell_text(align = "left", size = "medium"),
            cell_borders(sides = "all", color = "white")
        ),
        locations = cells_stub()
    ) %>% 
    tab_style(
        style = cell_text(
            align = "center", 
            size = "medium",
            weight = "normal",
            style = "normal"
        ),
        locations = list(
            cells_column_spanners(),
            cells_column_labels()
        )
    ) %>% 
    tab_style(
        style = cell_borders(sides = "all", color = "white"),
        locations = cells_body(columns = 1:9)
    ) %>% 
    tab_style(
        style = cell_borders(sides = "top", color = "white"),
        locations = cells_body(columns = 1:9)
    )
```

@@

```{r}
#| label: fig-stimuli-freq
stimuli %>% 
    distinct(group, word_1, .keep_all = TRUE) %>% 
    left_join(count(., group, name = "n_total")) %>% 
    mutate(test_language = paste0(group,  " (N = ", n_total, ")")) %>% 
    ggplot(aes(freq_zipf_2, fill = group)) +
    facet_wrap(~group, ncol = 1) +
    geom_histogram(bins = 25, colour = "white") +
    labs(
        x = "Lexical frequency (Zipf score)\nExtracted from CLEARPOND database", 
        y = "# trials", 
        fill = "Group"
    )  +
    scale_y_continuous(breaks = seq(0, 12, 2)) +
    theme(
        legend.position = "none",
        panel.grid.major.y = element_line(colour = "grey", linetype = "dotted")
    )
```

@@

```{r}
#| label: fig-model-rocs
model_rocs %>% 
    bind_rows(.id = "model") %>% 
    ggplot(aes(1-specificity, sensitivity, colour = model)) +
    facet_wrap(~model) +
    geom_line(aes(group = .draw), alpha = 0.1) +
    labs(x = "1 - Specificity", y = "Sensitivity", colour = "Model") + 
    scale_y_continuous(labels = percent) +
    scale_x_continuous(labels = percent) +
    theme_custom()

@@

```{r}
#| label: tbl-model-coefs
coefs <- model_parameters$fit_5 %>%
    as_tibble() %>% 
    clean_names() %>% 
    mutate(
        parameter = str_remove(parameter, "b_"),
        across(median:ci_high, ~ifelse(parameter=="(Intercept)", plogis(.), ./4))
    ) %>% 
    select(-ci)

str_repl <- c(
    "Intercept" = "Intercept",
    "freq_zipf_2_std" = "Frequency [+1 SD]",
    "nd_std" = "Neighbours [+1 SD]",
    "lv_std" = "Levenshtein [+1 SD]",
    "groupeng_spa" = "Group [ENG vs. SPA]",
    "groupeng_spa" = "Group [spa/cat-ENG vs. cat-SPA]",
    "groupcateng_spaeng" = "Group [cat-ENG vs. spa-ENG]",
    "nd_std:lv_std" = "Neigbours \u00d7 Cognateness"
)

@@

coefs %>% 
    mutate(
        parameter = factor(parameter, levels = names(str_repl), labels = str_repl),
        across(starts_with("ci_"), ~percent(., accuracy = 0.01)),
        ci = paste("[", ci_low, ", ", ci_high, "]")
    ) %>% 
    select(-starts_with("ci_"), -c(rhat, ess)) %>% 
    relocate(parameter, median, ci, log_bf) %>% 
    gt() %>% 
    fmt_percent(median) %>% 
    fmt_number(log_bf) %>% 
    cols_label(
        parameter = "Predictor",
        median = "Median",
        ci = "95% CrI",
        log_bf = "Log(Bayes Factor)"
    ) %>% 
    tab_source_note(tab_caption) %>% 
    cols_align(align = "right")
```

@@

```{r}
#| label: fig-questionnaire-model-rocs
questionnaire_model_rocs %>% 
    bind_rows(.id = "model") %>% 
    ggplot(aes(1-specificity, sensitivity, colour = model)) +
    facet_wrap(~model) +
    geom_line(aes(group = .draw), alpha = 0.1) +
    labs(x = "1 - Specificity", y = "Sensitivity", colour = "Model") + 
    scale_y_continuous(labels = percent) +
    scale_x_continuous(labels = percent) +
    theme_custom()

@@

## Confidence ratings and probability of correct responses

```{r}
#| label: fig-confidence-correct 
questionnaire_responses %>%
    mutate(lv = cut(lv, breaks = seq(0, 1, 0.1), include.lowest = TRUE)) %>% 
    group_by(word_1, group, lv) %>% 
    summarise(
        n = n(),
        confidence_mean = mean(confidence),
        knowledge_prop = mean(knowledge),
        correct = sum(correct),
        .groups = "drop"
    ) %>% 
    mutate(correct_prop = prop_adj(correct, n)) %>% 
    arrange(correct_prop) %>% 
    ggplot(aes(confidence_mean, correct_prop, label = word_1)) +
    facet_grid(group~lv) +
    geom_point(alpha = 0.5) +
    geom_text_repel() +
    labs(x = "Mean confidence score", y = "P(Correct)") +
    scale_y_continuous(labels = percent, breaks = seq(0, 1, 0.25), limits = c(0, 1))

```{r}
#| label: tbl-confidence-correct 
a <- questionnaire_responses %>%
    group_by(translation, group, lv) %>% 
    summarise(
        n = n(),
        confidence_mean = mean(confidence),
        knowledge_prop = mean(knowledge),
        correct_prop = mean(correct),
        .groups = "drop"
    ) %>% 
    arrange(group, lv) %>% 
    select(-n) %>% 
    gt(groupname_col = "group", rowname_col = "translation") %>% 
    tab_header(title = "Word-level accuracy scores",
               subtitle = "Observed proportion of correct translation of every word used in the questionnaire (only English participants translation Catalan or Spanish), with information about their Levenshtein similarity, and mean confidence of the translation and prior knowledge with the preented word reported by participants.") %>% 
    fmt_percent(c(lv, ends_with("prop"))) %>% 
    fmt_number(confidence_mean) %>% 
    cols_label(
        translation = "Translation (ordered by Levenshtein similarity)",
        lv = "Levenshtein similarity",
        confidence_mean = "Mean confidence (0-5)",
        knowledge_prop = "Knows the word (%)",
        correct_prop = "Accuracy (%)"
    ) %>% 
    tab_style(cell_text(weight = "bold"), cells_column_labels()) %>% 
    tab_style(list(cell_text(style = "italic", weight = "bold"), cell_fill("grey")), cells_row_groups()) %>% 
    tab_style(cell_text(align = "left"), cells_title(groups = "subtitle")) %>% 
    data_color(
        columns = correct_prop,
        colors = col_numeric(
            palette = rev(c("green", "yellow", "orange", "red")),
            domain = c(0, 1)
        )
    ) %>% 
    tab_footnote(
        footnote = "Presented word, correct translation, and their correspondent IPA phonological transcription",
        locations = cells_column_labels(columns = translation)
    ) %>% 
    tab_footnote(
        footnote = "Levenshtein distance between the IPA transcriptions of the presented word and its correct translation, divided by the number of characters of the longest transcription, and substracted from 1 to make it a similarity score.",
        locations = cells_column_labels(columns = lv)
    ) %>% 
    tab_footnote(
        footnote = "Mean of the reported confidence of participants' translations, in a scale from 0 to 5.",
        locations = cells_column_labels(columns = confidence_mean)
    ) %>% 
    tab_footnote(
        footnote = "Proportion of participants that reported knowing the translation of the word in Catalan or Spanish.",
        locations = cells_column_labels(columns = knowledge_prop)
    ) %>% 
    tab_footnote(
        footnote = "Proportion of participants that translated the presented word to English correctly, colour-coded with warm colours assiged to higher accuracy scores, and colder colours assigned to lower accuracy scores.",
        locations = cells_column_labels(columns = correct_prop)
    ) %>% 
    tab_stubhead(label = "Translation")

gtsave(a, "accuracy-confidence.html")
